---
title: "Exercise 2 - Fakebook Bus"
author: "Samuel"
date: "`r Sys.Date()`"
output: 
  github_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# References

<https://mr.schochastics.net/material/netvizr/>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(tidygraph)
library(igraph)
library(ggraph)

```

# Objective

The goal of this exercise is to find the optimal position to sit in the
bus.

We'll look at the network of the people in the bus.

# Creating the dataset

## Nodes

```{r}
nodes <- data.frame(Nodes = c(1,2,3,4,5,6,"A","B","C","D"))
nodes %>% head()

```

## Edges

For simplicity, those were created manually in Excel because it was
easier to do compared to in R.

```{r}
edges <- readxl::read_excel("ex2_edges.xlsx")
edges %>% head()

```

### Unidirectionnal edges

Doing that to remove the duplicates directions so the centrality
measures will be accurate. Counted twice otherwise.

```{r}
edges <- as_data_frame(simplify(graph_from_data_frame(edges, directed=FALSE)))

```

# Creating the graph

```{r}
graph <- tbl_graph(edges = edges, nodes=nodes, directed = FALSE)
autograph(graph)

```

# Visualize

Possible layouts: igraph_layouts \<- c('star', 'circle', 'gem', 'dh',
'graphopt', 'grid', 'mds', 'randomly', 'fr', 'kk', 'drl', 'lgl')

## Computing different parameters for the Network Graph

```{r}
# Clustering for node colors:
V(graph)$clu <- as.character(membership(cluster_louvain(graph)))

# Degree Centrality
V(graph)$degree <- degree(graph)

# Betweenness centrality
V(graph)$bet_centrality <- betweenness(graph)

```

Showing the degree of the vertex

```{r}
V(graph)$degree
```

Betweenes of the vertex

```{r}
V(graph)$bet_centrality

```

## Plot the results for Degree Centrality

```{r}
ggraph(graph, layout = "stress") +
  geom_edge_link(color = "gray66") +
  geom_node_point(
    aes(fill = clu, size = degree),  # size based on graph degree centrality
    shape = 21,
    ) +
  scale_fill_brewer(palette = "Dark2", na.value = "gray53") +
  geom_node_text(
    aes(label = Nodes),
    colour = "#000000",
    size = 5,
    repel = TRUE,
    family= "arial"
  ) +
  theme_graph() +
  theme(legend.position = "none",
        plot.title = element_text(size=12)) +
  labs(title="Fakebook Bus Network - Degree Centrality as Node Size")

```

***Analysis from this graph using degree centrality***

While the connection centrality clearly advantage B, C and D and
therefore the intern should choose between these three. From my point of
view, it would be advantageous for the intern to sit on B or C, because
it has the advantage of being connecting with the 3 groups clusters
defined with the Louvain method. Seat D is really far from the (A,2,1)
cluster.

## Plot the results for Betweenness Centrality

```{r}
ggraph(graph, layout = "stress") +
  geom_edge_link(color = "gray66") +
  geom_node_point(
    aes(fill = clu, size = bet_centrality*2),  # size based on nodes betweenness
    shape = 21,
    ) +
  scale_fill_brewer(palette = "Dark2", na.value = "gray53") +
  geom_node_text(
    aes(label = Nodes),
    colour = "#000000",
    size = 5,
    repel = TRUE,
    family= "arial"
  ) +
  theme_graph() +
  theme(legend.position = "none",
        plot.title = element_text(size=12)) +
  labs(title="Fakebook Bus Network - Betweenness Centrality as Node Size")



```

```{r include=FALSE}
V(graph)$bet_centrality

```

# Final conclusion 

Now, using the betweenness centrality, we can see that this gives an
advantage to seat A and B. Meaning that these two nodes are bridges to
the others from the shortest path. Therefore, based on the betweenness
metric and degree centrality, we should select seat B as it is slightly
better than seat C in terms of betweenness centrality.

In real life however, if I am the last one to sit, I prefer being at the
back of the bus so then you can kind of interact with the people at the
front and if I am the first one to sit, then it's better to be more in
front so then you can interact with people going to sit at the back!
